A flexible and intuitive analytic expression for the metallicity-dependent cosmic star formation rate

### 2022-06-27 src/tex/ms.tex
[showyourwork] overleaf sync

% Title
\title{ A convenient, flexible and intuitive analytic expression for the metallicity-dependent cosmic star formation rate  }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract with filler text
\begin{abstract}
New observational facilities are probing star formation and stellar transients and gravitational wave sources at ever increasing redshifts. Comparing these observations with predictions from model for stellar populations or the channels for the formation of gravitational wave sources require assumptions for the cosmic starformation rate density and, in particular, how this depends on metallicity.  Various approaches exist ranging from fully empirical prescriptions to prescriptions that are based on cosmological simulations. Both these approaches come with advantages and disadvantages and inherent uncertainties.

To make it easier to produce model predictions we propose a simple analytical prescription that captures the main behaviour of the advanced simulations. At the same time, the prescription comes with parameters that allow for variations in the overall shape and which can be easily interpreted. Our hope is that this expression provides a useful starting point for making predictions and comparison with observations. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction \label{sec: intro}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Since the delay time of binary black holes (BBHs) can range from Myrs to multiple Gyrs, the birth environment of BBHs varies significantly and we cannot ignore the cosmological (metallicity) evolution of the Universe. 
% The rate of Binary Black Hole (BBH) mergers at different redshifts depends both on the delay time distribution as well as the metallicity dependent star-formation rate density (\SFRDzZ).
% The assumed cosmic star formation rate density (SFRD) will therefore greatly influence the observed rate of double compact object mergers \citep{Neijssel+2019, Chruslinska2019_obs, Boco2020,Broekgaarden+2021a}. 
% Moreover, it is the nuclear reactions in massive stars are responsible for the metal enrichment of the Universe at high redshift. The birth of massive stars, and their death as BHs are thus inseparable from the evolution of the metallicity distribution of the Universe.

% %%  metallicity density distribution %%
% Merging BBHs are found to form more efficiently at low metallicities e.g. \citep[e.g.][]{BelczynskiVink2010, Stevenson+2017,Mapelli2017,Chruslinska2019_effectCO}. Therefore the assumed metallicity specific star formation rate will in particular have a large impact on the inferred merger rate \citep[e.g.][]{Chruslinska2019_obs,Chruslinska2019_effectCO, Neijssel2019, Broekgaarden+2021a}.

A myriad of astrophysical phenomena depend critically on the rate of star formation throughout the cosmic history of the Universe. 
Transient phenomena, including core collapse supernovae, (pulsational) pair-instability supernovae, long gamma-ray bursts and gravitational wave (GW) events are particularly sensitive to the metallicity at which star formation occurs at different epochs throughout the Universe.
In particular, the field of gravitational astronomy has seen explosive growth in the past decade \citep[][]{GWTC1,GWTC2,GWTC3}. In order to correctly model and interpret these observations, it is fundamental to know rate of star formation at different metallicities throughout cosmic history, i.e. the metallicity-dependent star formation rate density \SFRDzZ \citep[see also the recent review by][]{chruslinska2022_review}.

@@

\subsection{The overall cosmic star formation rate density}
For the star formation rate density, we assume the analytical form proposed by \cite{MadauDickinson2014},
\begin{eqnarray}
\label{eq: sfr1}
\boxed{
    \SFRDz  = 
    \frac{d^2 M_{\rm SFR}}{dt dV_c} (z)= 
    a \frac{\left(1 + z\right)^b}{1 + \left[ (1 + z)/c \right]^d} 
    }\,
\end{eqnarray}
in units of $\left[ \Msun \,yr^{-1} \,cMpc^{-3} \right]$. This introduces four parameters, $a$ which sets the overal normalisation and which has the same units as \SFRDz and $b,c$ and $d$ which are unitless and which govern the shape of the overal cosmic starformation rate density with redshift. \\

% \subsection{Combined function}
Lastly, we combine equations \ref{eq: z log skew} and \ref{eq: sfr1} to form a full metallicity specific starformation rate density as described in equation \ref{eq: total sfrd}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\script{FitComparison_3panelPlot.py}
\includegraphics[width=\textwidth]{figures/SFRD_FIT_evaluation_compare.pdf}
\caption{Our fiducial \SFRDzZ model, adopting the best fitting parameters (listed on the top right) to fit the TNG100 simulations.
The top panel shows the full two dimensional \SFRDzZ linear in time. The bottom left (right) panel shows slices of the distribution in redshift (metallicity). Each slice is displaced by 0.01$\Msun \yr^{-1}\Mpc^{-3}$. We show the TNG100 simulation data with thick gray lines. 
For comparison, we also show the phenomenological model from \protect\cite{Neijssel+2019} in each panel with grey dotted lines. For the latter, the contours in the top panel range from $10^{-7} - 10^{-2} \Msun \yr^{-1}\Mpc^{-3}$. This shows that our analytical model adequately captures the \SFRDzZ of the TNG100 simulations.
 \label{fig: fit SFRD}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We fit our newly defined functional form of \SFRDzZ as defined by equations \ref{eq: total sfrd}, \ref{eq: z log skew} and \ref{eq: sfr1} to the Illustris-TNG cosmological simulations. 
We fit for the following nine free parameters $\alpha, \mu_0, \mu_z, \omega_0, \omega_z$, which govern the metallicity dependence and $a,b, c$ and $d$, which set the overall starformation rate.

@@

% %
% The task we have at hand here is not a standard textbook fitting problem given the specific aims and requirements that we have that are driven by the astrophysical applications we have in mind. Qualitatively, our aims are the following. Our main aim is to find solution for the free parameters, for which the expression reproduces the overall behaviour that is observed in the cosmological simulations. For the purposes we have in mind, it will be more important to prioritise fitting the large scale trends, while we are not so interested in smaller scale fluctuations. We further prioritise getting the bulk of the cosmic star formation right and are less interested in capturing the behaviour in metallicity and redshift bins where the cosmic starformation rate is very low and thus insignificant for the overall picture. Finally, we are especially interested in capturing the asymmetry of the metallicity distribution. 
% There is not one unique procedure that will full fill these criteria.  After experimenting with different criteria we have adopted the procedure below, which we believe works well for the aims we have in mind.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Choices and binning of the data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We fit equation \ref{eq: total sfrd} to the metallicity-dependent starformation rate of the starforming gas in the TNG100 simulation. For this we use a binned version of the TNG data $\SFRDzZ_{\rm sim}$. We consider metallicities between $Z= -5$ to $\log_{10} Z= 0$ in 30 bins, where we use $Z_i$ to refer to the logarithmic centers of the bins. We ignore star formation in metallicities $\log_{10} Z \le -5$ as this accounts for less than 1\% of the total cosmic starformation rate in these simulations.

We consider bins in redshifts between $z=0$ and $z=10$, with a step size of $dz=0.05$, where $z_j$ refers to the centers of the bins.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Optimisation function}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To find a solution we use a method based on the sum of the quadratic differences between the simulations and our fit function. Using a vanilla $\chi$-squared approach does not serve our purposes very well as it does a poor job in fitting regions where the starformation is very low.  Using a $\chi$-squared approach on the logarithm of the function instead places far too much weight on trying to fit the starformation rate in regions where the rate is very low or not even significant.  After experimenting, we find that the following approach gives us satisfactory results.

We first consider a given redshift $z_j$.  For this redshift we compute the sum of the squared residuals between the cosmological simulation and our fit.  
%
\begin{equation}
\label{eq: chisquare}
    \chi^2 (z_j) \equiv \sum_{Z_i} \left( 
        \mathcal{S}(Z_i,z_j)_{\rm sim} - 
        \mathcal{S}(Z_i,z_j)_{\rm fit}\right)^2
    %chi_square = ((obs - model )**2)/np.sum(model) 
\end{equation}
%
Here, the variable $Z_i$ runs over all redshift bins, but excludes bins where the starformation rate density is lower $10^{-10} \Msun \yr \Mpc^{-3}$ per bin.

Subsequently we sum the $\chi^2 (z_j)$ for all redshift bins $z_j$. To ensure that our fit procedure gives sufficient weight to the behaviour at all redshifts, we find that we need to introduce a penalisation factor to somewhat reduce the contribution of redshifts where the peak of cosmic starformation occurs, while increasing the weight where at redshifts where the overal cosmic starformation rate is lower.  To achieve this we divide $\chi^2 (z_j)$ by the starformation $\sum_{Z_i} \mathcal{S}(Z_i,z_j)$ per redshift bin before adding the contribution of all redshifts.  Our final expression for the cost function reads

@@

In Figure \ref{fig: fit SFRD} we show our fiducial \SFRDzZ model at different redshifts and metallicities. For clarity, 
In general, our analytical model captures the metallicity dependent star formation in the TNG100 simulations well (bottom panels). 
The skewed-log normal metallicity distribution is able to reproduce the overall behaviour that is observed in TNG100 \citep[bottom left panel, but cf. ][for an in-depth discussion of low metallicity star formation in the TNG50 simulation]{Pakmor+2022}
Only minor features like the additional bump just above $\log_{10}(Z) = -2$ at redshift 2 are missed. 
However, for our purposes, it is more important to prioritise fitting the large scale trends, while we are not so interested in smaller scale fluctuations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\script{SFR_z.py}
\includegraphics[width=0.49\textwidth]{figures/SFR_redshift.pdf}
\caption{Comparison of different \SFRDz
 \label{fig: SFR(z)}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We consider variations of each of the parameters that determine the shape of the cosmic metallicity distribution \dpdZ (left two columns of Table \ref{tab: fit params}) and two variations of the \SFRDz, where we vary the complete shape of the \SFRDz (i.e. all four parameters) at once to match the values used in \cite{Madau+2017} and \cite{Neijssel2019} respectively. To determine the range of parameters that is reasonably allowed by observations, we compare the \SFRDzZ in the metallicity-redshift plane to the high and low metallicity extreme from \cite{Chruslinska+2021} by eye for every variation. \\

@@

% The normalization at z=0 changes the overal rate / normalization
We find that parameter variations that affect shape of \SFRDzZ at low redshift primarily change the normalisation of the mass distribution (see also  $\mathcal{R}_{0.2}$, the total BBH merger rate at redshift 0.2, annotated in the legends of Figure \ref{fig: mass dists}). 
These variations include the width of the cosmic metallicity distribution at $z=0$, $\omega_0$, the mean metallicity of the cosmic metallicity distribution at $z=0$, $\mu_0$, and the skewness of the cosmic metallicity distribution, $\alpha$ (left column of Figure \ref{fig: mass dists}).
Model variations that increase the amount of star formation at low metallicity (i.e. for a low mean metallicity $\mu_0=0.015$ and a wide metallicity distribution $\omega_0 = 1.4$ ) increase the predicted BBH merger rate. This is consistent with other work that finds merging BBHs form more efficiently at low metallicities \citep[e.g.][]{BelczynskiVink2010, Stevenson+2017,Mapelli2017,Chruslinska2019_effectCO,Broekgaarden+2021b}.

@@

% Effect of the star formation 
Lastly, changing the \SFRDz has a relatively small effect. 
The \SFRDz from \cite{Madau+2017} and \cite{Neijssel+2019} both peak at approximately redshift 2. Our fiducial model (that follows the \SFRDz from TNG) peaks at about redshift 3. The overall normalization of \cite{Neijssel+2019} is higher than the normalization of our fiducial model. Regardless of these differences, the rate of BBH mergers at $z=0.2$ changes no more than 20\%, and the shape of the primary mass distribution remains almost constant.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary \label{sec: summary}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We present a flexible analytic expression for the metallicity-dependent starformation rate, \SFRDzZ (equations \ref{eq: total sfrd}, \ref{eq: z log skew} and \ref{eq: sfr1}). An analytical expression allows for controlled experiments of the effect of \SFRDzZ on dependent rates, such as the rate and mass distribution of merging BBHs. 
The novelty of the model presented in this work is that it adopts a skewed-lognormal for the distribution of metallicities at every redshift (\dpdZ).

We fit our analytical expression for \SFRDzZ to the starforming-gas in the TNG100 simulation, and provide the best fit parameters in Table \ref{tab: fit params}. We show that our model captures the shape and general behaviour of the cosmological simulations well (Figure \ref{fig: fit SFRD}).

As an example, we apply our model to calculate the local rate and mass distribution of the more massive components from merging BBHs (\Mbheen) in Figure \ref{fig: mass dists}). 
We systematically vary all five parameters ($\alpha, \mu_0, \mu_z, \omega_0 and \omega_z$) that shape the cosmic metallicity distribution \dpdZ, and explore two addition variations of \SFRDz \citep[following][respectively]{Madau+2017,Neijssel+2019}.

@@

The flexibility of the model presented in this work can capture the large uncertainties that remain in the shape and normalisation of the metallicity dependent cosmic star formation history. As such, we anticipate that it will be a helpful tool in other works that present \SFRDzZ dependent rates.

>>>>

% Title
\title{ A flexible and intuitive analytic expression for the metallicity-dependent cosmic star formation rate  }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract with filler text
\begin{abstract}
New observational facilities are probing astrophysical transients such as gravitational wave (GW) sources at ever increasing redshifts. To interpret these observation we need to compare them to predictions from stellar population models. These models are very sensitive to the metallicity-dependent star formation rate density (\SFRDzZ). 
% Large uncertainties remain in the value and shape of the \SFRDzZ, especially at the low metallicities that are crucial to the formation of GW events.
Understanding the behaviour of the adopted \SFRDzZ is thus essential to explain simulation results. 
%
% Various approaches to approximate \SFRDzZ exist, ranging from fully empirical prescriptions to prescriptions that are based on cosmological simulations.
In this work, we propose a simple analytical form for \SFRDzZ. The novelty of this model lies in a skewed-lognormal distribution of metallicities, that captures the tail of low metallicity star formation at low redshift. 
Variations of this analytical form can be easily interpreted, because the parameters are intuitively linked to the shape of \SFRDzZ. %Our hope is that this expression provides a useful starting point for making predictions and comparison with observations.
%
We fit our analytical form to the starforming gas of the cosmological TNG100 simulations and find our model is able to capture the main behaviour. 
%
As an example, we investigate the effect of systematic variations in the \SFRDzZ parameters on the predicted mass distribution of locally merging binary black holes (BBH). Our main findings are I) the low mass end is least affected by variations in \SFRDzZ, II) the distribution of metallicities at low redshift affects the normalisation of of the BBH merger rate, while III) the redshift evolution of the metallicity distribution affects the slope of the mass distribution beyond the peak value. 
\end{abstract}

% Interpreting these observations requires comparison to predictions from stellar population models. However, the model predictions  are very sensitive to the adopted assumptions about the metallicity-dependent star formation rate density. Hence in order to 
% Comparing these observations to predictions from stellar population models requires assumptions about the metallicity-dependent star formation rate density. Various approaches exist 
% ranging from fully empirical prescriptions to prescriptions that are based on cosmological simulations. Both these approaches come with advantages and disadvantages and inherent uncertainties.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction \label{sec: intro}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A myriad of astrophysical phenomena depend critically on the rate of star formation throughout the cosmic history of the Universe. Transient phenomena, including core collapse supernovae, (pulsational) pair-instability supernovae, long gamma-ray bursts and gravitational wave (GW) events are particularly sensitive to the metallicity at which star formation occurs at different epochs throughout the Universe.
In particular, the field of gravitational astronomy has seen explosive growth in the past decade \citep[][]{GWTC1,GWTC2,GWTC3}. In order to correctly model and interpret these observations, it is fundamental to know rate of star formation at different metallicities throughout cosmic history, i.e. the metallicity-dependent star formation rate density \SFRDzZ \citep[see also the recent review by][]{chruslinska2022_review}.

@@

\subsection{The overall cosmic star formation rate density}
For the star formation rate density, we assume the analytical form proposed by \cite{MadauDickinson2014},
\begin{eqnarray}
\label{eq: sfr1}
\boxed{
    \SFRDz  = 
    \frac{d^2 M_{\rm SFR}}{dt dV_c} (z)= 
    a \frac{\left(1 + z\right)^b}{1 + \left[ (1 + z)/c \right]^d} 
    }\,
\end{eqnarray}
in units of $\left[ \Msun \,yr^{-1} \,cMpc^{-3} \right]$. This introduces four parameters, $a$ which sets the overal normalisation and which has the same units as \SFRDz and $b,c$ and $d$ which are unitless and which govern the shape of the overal cosmic star formation rate density with redshift. \\

% \subsection{Combined function}
Lastly, we combine equations \ref{eq: z log skew} and \ref{eq: sfr1} to form a full metallicity specific star formation rate density as described in equation \ref{eq: total sfrd}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\script{FitComparison_3panelPlot.py}
\includegraphics[width=0.9\textwidth]{figures/SFRD_FIT_evaluation_compare.pdf}
\caption{Our fiducial \SFRDzZ model, adopting the best fitting parameters (listed on the top right) to fit the TNG100 simulations.
The top panel shows the full two dimensional \SFRDzZ linear in time. The bottom left (right) panel shows slices of the distribution in redshift (metallicity). Each slice is displaced by 0.01$\Msun \yr^{-1}\Mpc^{-3}$. We show the TNG100 simulation data with thick gray lines. 
For comparison, we also show the phenomenological model from \protect\cite{Neijssel+2019} in each panel with grey dotted lines. For the latter, the contours in the top panel range from $10^{-7} - 10^{-2} \Msun \yr^{-1}\Mpc^{-3}$. This shows that our analytical model adequately captures the \SFRDzZ of the TNG100 simulations.
 \label{fig: fit SFRD}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We fit our newly defined functional form of \SFRDzZ as defined by equations \ref{eq: total sfrd}, \ref{eq: z log skew} and \ref{eq: sfr1} to the Illustris-TNG cosmological simulations. 
We fit for the following nine free parameters $\alpha, \mu_0, \mu_z, \omega_0, \omega_z$, which govern the metallicity dependence and $a,b, c$ and $d$, which set the overall star formation rate.

@@

% %
% The task we have at hand here is not a standard textbook fitting problem given the specific aims and requirements that we have that are driven by the astrophysical applications we have in mind. Qualitatively, our aims are the following. Our main aim is to find solution for the free parameters, for which the expression reproduces the overall behaviour that is observed in the cosmological simulations. For the purposes we have in mind, it will be more important to prioritise fitting the large scale trends, while we are not so interested in smaller scale fluctuations. We further prioritise getting the bulk of the cosmic star formation right and are less interested in capturing the behaviour in metallicity and redshift bins where the cosmic star formation rate is very low and thus insignificant for the overall picture. Finally, we are especially interested in capturing the asymmetry of the metallicity distribution. 
% There is not one unique procedure that will full fill these criteria.  After experimenting with different criteria we have adopted the procedure below, which we believe works well for the aims we have in mind.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Choices and binning of the data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We fit equation \ref{eq: total sfrd} to the metallicity-dependent star formation rate of the starforming gas in the TNG100 simulation. For this we use a binned version of the TNG data $\SFRDzZ_{\rm sim}$. We consider metallicities between $Z= -5$ to $\log_{10} Z= 0$ in 30 bins, where we use $Z_i$ to refer to the logarithmic centers of the bins. We ignore star formation in metallicities $\log_{10} Z \le -5$ as this accounts for less than 1\% of the total cosmic star formation rate in these simulations.
We consider bins in redshifts between $z=0$ and $z=10$, with a step size of $dz=0.05$, where $z_j$ refers to the centers of the bins.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Optimisation function}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To find a solution we use a method based on the sum of the quadratic differences between the simulations and our fit function. Using a vanilla $\chi$-squared approach does not serve our purposes very well as it does a poor job in fitting regions where the star formation is very low.  Using a $\chi$-squared approach on the logarithm of the function instead places far too much weight on trying to fit the star formation rate in regions where the rate is very low or not even significant.  After experimenting, we find that the following approach gives us satisfactory results.

We first consider a given redshift $z_j$.  For this redshift we compute the sum of the squared residuals between the cosmological simulation and our fit.  
%
\begin{equation}
\label{eq: chisquare}
    \chi^2 (z_j) \equiv \sum_{Z_i} \left( 
        \mathcal{S}(Z_i,z_j)_{\rm sim} - 
        \mathcal{S}(Z_i,z_j)_{\rm fit}\right)^2
    %chi_square = ((obs - model )**2)/np.sum(model) 
\end{equation}
%
Here, the variable $Z_i$ runs over all redshift bins, but excludes bins where the star formation rate density is lower $10^{-10} \Msun \yr \Mpc^{-3}$ per bin.

Subsequently we sum the $\chi^2 (z_j)$ for all redshift bins $z_j$. To ensure that our fit procedure gives sufficient weight to the behaviour at all redshifts, we find that we need to introduce a penalisation factor to somewhat reduce the contribution of redshifts where the peak of cosmic star formation occurs, while increasing the weight where at redshifts where the overal cosmic star formation rate is lower.  To achieve this we divide $\chi^2 (z_j)$ by the star formation $\sum_{Z_i} \mathcal{S}(Z_i,z_j)$ per redshift bin before adding the contribution of all redshifts.  Our final expression for the cost function reads

@@

In Figure \ref{fig: fit SFRD} we show our fiducial \SFRDzZ model at different redshifts and metallicities. For clarity, we also show the overal rate of star formation \SFRDz in Figure \ref{fig: SFR(z)}.
In general, our analytical model captures the metallicity dependent star formation in the TNG100 simulations well (bottom panels of Figure \ref{fig: fit SFRD}). 
The skewed-log normal metallicity distribution is able to reproduce the overall behaviour that is observed in TNG100 \citep[bottom left panel, but cf. ][for an in-depth discussion of low metallicity star formation in the TNG50 simulation]{Pakmor+2022}
Only minor features like the additional bump just above $\log_{10}(Z) = -2$ at redshift 2 are missed. 
However, for our purposes, it is more important to prioritise fitting the large scale trends, while we are not so interested in smaller scale fluctuations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\script{SFR_z.py}
\includegraphics[width=0.47\textwidth]{figures/SFR_redshift.pdf}
\caption{Comparison of different \SFRDz
 \label{fig: SFR(z)}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We consider variations of each of the parameters that determine the shape of the cosmic metallicity distribution \dpdZ (left two columns of Table \ref{tab: fit params}) and two variations of the \SFRDz, where we keep the metallicity distribution \dpdZ fixed, but we vary the complete shape of the \SFRDz (i.e. all four parameters) at once to match the values used in \cite{Madau+2017} and \cite{Neijssel2019} respectively (see Figure \ref{fig: SFR(z)} for reference). To determine the range of parameters that is reasonably allowed by observations, we compare the \SFRDzZ in the metallicity-redshift plane to the high and low metallicity extreme from \cite{Chruslinska+2021} by eye for every variation. \\

@@

% The normalisation at z=0 changes the overal rate / normalisation
We find that parameter variations that affect shape of \SFRDzZ at low redshift primarily change the normalisation of the mass distribution (see also  $\mathcal{R}_{0.2}$, the total BBH merger rate at redshift 0.2, annotated in the legends of Figure \ref{fig: mass dists}). 
These variations include the width of the cosmic metallicity distribution at $z=0$, $\omega_0$, the mean metallicity of the cosmic metallicity distribution at $z=0$, $\mu_0$, and the skewness of the cosmic metallicity distribution, $\alpha$ (left column of Figure \ref{fig: mass dists}).
Model variations that increase the amount of star formation at low metallicity (i.e. for a low mean metallicity $\mu_0=0.015$ and a wide metallicity distribution $\omega_0 = 1.4$ ) increase the predicted BBH merger rate. This is consistent with other work that finds merging BBHs form more efficiently at low metallicities \citep[e.g.][]{BelczynskiVink2010, Stevenson+2017,Mapelli2017,Chruslinska2019_effectCO,Broekgaarden+2021b}.

@@

% Effect of the star formation 
Lastly, changing the \SFRDz has a relatively small effect. We vary 
The \SFRDz from \cite{Madau+2017} and \cite{Neijssel+2019} both peak at approximately redshift 2. Our fiducial model (that follows the \SFRDz from TNG) peaks at about redshift 3. The overall normalisation of \cite{Neijssel+2019} is higher than the normalisation of our fiducial model. Regardless of these differences, the rate of BBH mergers at $z=0.2$ changes no more than 20\%, and the shape of the primary mass distribution remains almost constant.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary \label{sec: summary}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We present a flexible analytic expression for the metallicity-dependent star formation rate, \SFRDzZ (equations \ref{eq: total sfrd}, \ref{eq: z log skew} and \ref{eq: sfr1}). An analytical expression allows for controlled experiments of the effect of \SFRDzZ on dependent rates, such as the rate and mass distribution of merging BBHs. 
The novelty of the model presented in this work is that it adopts a skewed-lognormal for the distribution of metallicities at every redshift (\dpdZ).

We fit our analytical expression for \SFRDzZ to the star-forming gas in the TNG100 simulation, and provide the best fit parameters in Table \ref{tab: fit params}. We show that our model captures the shape and general behaviour of the cosmological simulations well (Figure \ref{fig: fit SFRD}).

As an example, we apply our model to calculate the local rate and mass distribution of the more massive components from merging BBHs (\Mbheen) in Figure \ref{fig: mass dists}). 
We systematically vary all five parameters ($\alpha, \mu_0, \mu_z, \omega_0$ and $\omega_z$) that shape the cosmic metallicity distribution \dpdZ, and explore two addition variations of \SFRDz, following \cite{Madau+2017} and \cite{Neijssel+2019} respectively.

@@

The flexibility of the model presented in this work can capture the large uncertainties that remain in the shape and normalisation of the metallicity dependent cosmic star formation history. 
Our hope is that this expression will provide a useful starting point for making predictions and comparison with observations.
% As such, we anticipate that it will be a helpful tool in other works that present \SFRDzZ dependent rates.
